---
output: 
  html_document:
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: sentence
---

# Introduction

We aim to determine if there is a difference between biological (bAge) and chronological (cAge) age in Turner syndrome (45,X) and Klinefelter syndrome (47,XXY) compared to euploid controls (46,XX; 46,XY).
For this, we use the MethylClock and DunedinPACE R-packages as well as the GrimAge calculater.

# Set up environment

```{r}

path = "/path/to/folder/"
path_data = "/path/to/folder/DATA/"
path_plots = "/path/to/folder/PLOTS/"

# We load the required packages. 
library(tidyverse)
library(methylclock)
library(ggplot2)
library(ggpubr)
library(minfi)
library(generics)
library(Hmisc)
library(emmeans)
library(DunedinPACE)

col = c("TS" = "#F4A261",
        "Female" ="#922D50", 
        "Male" = "#477a8f",
        "KS" = "#3fc49c")

order <- c("TS","Female","KS","Male")

```

# Load data

We load the following datasets:

-   KS 450K (47,XXY; 46,XY) beta-values

```{r KS 450K data, message=FALSE, warning=FALSE, results = "hide", eval = FALSE}

KS450K <- readRDS(paste0(path,"path/methylation_data.rds")) 
KS450K_info <- readRDS(paste0(path,"path/clinical_traits.rds"))

# Make sure that samples are in the same order:
identical(colnames(KS450K),
          KS450K_info$slide_array)

# If not, use this: 
KS450K <- KS450K[, KS450K_info$slide_array]

# The order should be correct now:
identical(colnames(KS450K),
          KS450K_info$slide_array)

```

-   TS 450K (45,X + mosaics; 46,XX) beta-values

```{r TS 450K data, message=FALSE, warning=FALSE, results = "hide", eval = FALSE}
TS450K <- readRDS(paste0(path,"path/methylation_data.rds")) 
TS450K_info <- readRDS(paste0(path,"path/clinical_traits.rds"))

# Make sure that samples are in the same order:
identical(colnames(TS450K),
          TS450K_info$ID_new)
```

We combine the two datasets:

```{r combined data, message=FALSE, warning=FALSE, results = "hide", eval = FALSE}
CGs_KS450K <- rownames(KS450K)
CGs_TS450K <- rownames(TS450K)

CGs <- intersect(CGs_KS450K,CGs_TS450K)

M <- cbind(KS450K[CGs,], TS450K[CGs,])
```

Generate sample info:

```{r message=FALSE, warning=FALSE, results = "hide", eval = FALSE}
info <- rbind(KS450K_info,S450K_info ) %>% as.data.frame()

identical(colnames(M),
          info$Sample_Name)
```

# Run algorithms

## GrimAge

Load GrimAge data:

```{r}
GrimAge_res <- read_csv(paste0(path_data, "RESULTS/DNAmAgeCalcProject_Results.csv")) %>% 
  mutate(SID = substr(SID, 2, nchar(SID)))
GrimAge_dict <- read_csv(paste0(path_data, "RESULTS/DNAmAgeCalcProject_DataDict.csv"))
```

## Run MethylClock

We run Methylclock with selected clocks.

```{r run, message=FALSE, warning=FALSE, eval=FALSE}
age <- DNAmAge(M,
               age=info$age,
               cell.count=TRUE, 
               clocks = c("Horvath", "Hannum", "BLUP", "EN", "TL"))
```

## DunedinPACE

Run DunedinPACE:

```{r message=FALSE, warning=FALSE}
PACE <- PACEProjector(M) 
```

## Combined data frames

We make a dataframe from all algorithms.

```{r message=FALSE, warning=FALSE, eval = FALSE}

df <- merge(
  
        merge(
    
          merge(age, 
                as.data.frame(PACE) %>% 
                mutate(id = rownames(.)), 
                by = "id") %>% 
            mutate(id = gsub("-",".", id),
                   id = gsub(" ",".", id)), 
                                        
              GrimAge_res %>% select(SID, 
                                     "Body Mass Index", 
                                     "Body Fat %", 
                                     "HDL Cholesterol", 
                                     "Waist:Hip Ratio", 
                                     predictedGender, 
                                     DNAmGrimAge2BasedOnRealAge, 
                                     DNAmGrimAge2BasedOnPredictedAge, 
                                     AgeAccelGrim, 
                                     DNAmPhenoAge, 
                                     AgeAccelPheno, 
                                     IEAA, 
                                     IEAA.Hannum, 
                                     EEAA, 
                                     "Epigenetic Age (Zhang)", 
                                     AgeAccelerationResidual, 
                                     AgeAccelerationResidualHannum, 
                                     DNAmAgeHannum, 
                                     DNAmAge, 
                                     DNAmAgeSkinBloodClock), 
              by.x = "id", by.y = "SID"),
        
            info %>% select(-age) %>% 
              mutate(Sample_Name = gsub("-",".", Sample_Name),
                     Sample_Name = gsub(" ",".", Sample_Name)), 
        by.x = "id", by.y = "Sample_Name") %>% 
  rename("age" = "cAge")

```

## Save data

```{r message=FALSE, warning=FALSE, eval=FALSE}
save(age, 
     PACE, 
     df, 
     file = paste0(path_data,"age_predictions.Rdata"))
```

# Data analysis

## Figure 1

Set data frames for DNAmAge and AgeAcc:

```{r}
age_df <- df %>% select("BLUP", "EN", "Hannum", "Horvath", group, cohort, cAge)

# Set data frame for TS cohort
DNAmAge_TS <- age_df %>% 
  filter(cohort == "TS cohort") 

DNAmAge_mean_TS <- age_df %>% 
  filter(cohort == "TS cohort")  %>% 
  mutate(DNAmAge = rowMeans(select(DNAmAge_TS, BLUP, EN, Hannum, Horvath))) %>% 
  select(DNAmAge, cAge, group)

# Set data frame for KS cohort
DNAmAge_KS <- age_df %>% 
                filter(cohort == "KS cohort") 

DNAmAge_mean_KS <- age_df %>% 
  filter(cohort == "KS cohort")  %>% 
  mutate(DNAmAge = rowMeans(select(DNAmAge_KS, BLUP, EN, Hannum, Horvath))) %>% 
  select(DNAmAge, cAge, group)
```

```{r}
acc_df <- df %>% select(colnames(df)[grepl("ageAcc", colnames(df))],
                        cAge, group, cohort) %>% 
  pivot_longer(cols = colnames(df)[grepl("ageAcc", colnames(df))]) %>% 
  mutate(accType = sub("\\..*", "", name),
         Algorithm = sub(".*\\.", "", name))

acc_df_mean <- df %>% 
  mutate(ageAcc = rowMeans(select(df,ageAcc.BLUP, ageAcc.EN, ageAcc.Hannum, ageAcc.Horvath)),
         ageAcc2 = rowMeans(select(df,ageAcc2.BLUP, ageAcc2.EN, ageAcc2.Hannum, ageAcc2.Horvath)),
         ageAcc3 = rowMeans(select(df,ageAcc3.BLUP, ageAcc3.EN, ageAcc3.Hannum, ageAcc3.Horvath))) %>% 
  select(ageAcc, ageAcc2, ageAcc3, cohort, group)
```

### DNAmAge

```{r}
# Test on means across the four clocks 
emm <- emmeans(lm(DNAmAge ~ cAge + group, data=DNAmAge_mean_TS), pairwise ~ group)
p.value <- as.data.frame(emm$contrasts)$p.value

merge(emm$emmeans,
      DNAmAge_mean_TS) %>% 
ggplot() +
  geom_jitter(aes(x = factor(group, levels = order), y = DNAmAge, fill = group, col = group), 
              width = .4,alpha = .4) +
  geom_errorbar(aes(x = group, y = emmean, ymin = lower.CL, ymax = upper.CL), 
               stat = "identity", width = .35, color = "black") +
    geom_point(aes(x = group, y = emmean, fill = group),
             size = 2, alpha = 1, shape = 21, color = "black") + 
  annotate("text", x = 1.5, y = 80, 
                label = paste("p =",round(p.value, digits = 4)), size = 3, 
            vjust = -0.5) +
  theme_bw() + theme(legend.position = "none") +
  scale_color_manual(values = col) + 
  scale_fill_manual(values = col) +
  labs(x = "", y = "DNAmAge") +
  ylim(15,85) -> DNAmAge_mean_p_TS

summary(lm(DNAmAge ~ cAge * factor(group, levels = c("Female","TS")), 
           data = DNAmAge_mean_TS)) %>% coef()

DNAmAge_mean_TS %>% 
  ggplot(aes(cAge, DNAmAge, color = factor(group, levels = order), fill = factor(group, levels = order))) +
  geom_abline(slope = 1, color = "grey80", linetype = "dashed") + 
  geom_point(alpha = .4) +
  geom_line(stat="smooth",method = "lm", alpha = 1, linewidth = .75) +
  theme_bw()+
  theme(legend.position = "none") +
  scale_color_manual(values = col) + 
  scale_fill_manual(values = col) +
  labs(x = "cAge", y = "DNAmAge") +
  ylim(15,85) -> DNAmAge_mean_p_TS2

```

```{r}
# Test on means across the four clocks 
emm <- emmeans(lm(DNAmAge ~ cAge + group, data=DNAmAge_mean_KS), pairwise ~ group)
p.value <- as.data.frame(emm$contrasts)$p.value

merge(emm$emmeans,
      DNAmAge_mean_KS) %>% 
ggplot() +
  geom_jitter(aes(x = factor(group, levels = order), y = DNAmAge, fill = group, col = group), 
              width = .4,alpha = .4) +
  geom_errorbar(aes(x = group, y = emmean, ymin = lower.CL, ymax = upper.CL), 
               stat = "identity", width = .35, color = "black") +
    geom_point(aes(x = group, y = emmean, fill = group),
             size = 2, alpha = 1, shape = 21, color = "black") + 
  annotate("text", x = 1.5, y = 70, 
                label = paste("p =",round(p.value, digits = 3)), size = 3, 
            vjust = -0.5) +
  theme_bw() + theme(legend.position = "none") +
  scale_color_manual(values = col) + 
  scale_fill_manual(values = col) +
  labs(x = "", y = "DNAmAge") +
  ylim(15,75) -> DNAmAge_mean_p_KS

summary(lm(DNAmAge ~ cAge * factor(group, levels = c("Male","KS")), 
           data = DNAmAge_mean_KS)) %>% coef()

DNAmAge_mean_KS %>% 
  ggplot(aes(cAge, DNAmAge, color = factor(group, levels = order), fill = factor(group, levels = order))) +
  geom_abline(slope = 1, color = "grey80", linetype = "dashed") + 
  geom_point(alpha = .4) +
  geom_line(stat="smooth",method = "lm", alpha = 1, linewidth = .75) +
  theme_bw()+
  theme(legend.position = "none") +
  scale_color_manual(values = col) + 
  scale_fill_manual(values = col) +
  labs(x = "cAge", y = "DNAmAge") +
  ylim(15,75) -> DNAmAge_mean_p_KS2
```

### Acceleration

```{r}

#TS
emm <- emmeans(lm(ageAcc ~ group, data=acc_df_mean %>% filter(cohort == "TS cohort")), pairwise ~ group)
p.value <- as.data.frame(emm$contrasts)$p.value

merge(emm$emmeans,acc_df_mean %>% filter(cohort == "TS cohort")) %>% 
ggplot(aes(group, ageAcc, color = group, fill = group)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey80")+
  geom_point(alpha = .4, position = position_jitter(width = .35)) +
  geom_errorbar(aes(x = group, y = emmean, ymin = lower.CL, ymax = upper.CL), 
               stat = "identity", width = .35, color = "black") +
    geom_point(aes(x = group, y = emmean, fill = group),
             size = 2, alpha = 1, shape = 21, color = "black") + 
  annotate("text", x = 1.5, y = 15, label = paste("p =", signif(p.value, 2)), size = 3, 
            vjust = -0.5, color = "black") +
  theme_bw() + theme(legend.position = "none",
                     plot.title = element_text(hjust = .5))+
  scale_color_manual(values = col) + 
  scale_fill_manual(values = col) +
  labs(x = "", y = "Acceleration (years)", title = "ageAcc") +
  ylim(-8,17) -> TSa1

emm <- emmeans(lm(ageAcc2 ~ group, data=acc_df_mean %>% filter(cohort == "TS cohort")), pairwise ~ group)
p.value <- as.data.frame(emm$contrasts)$p.value

merge(emm$emmeans,acc_df_mean %>% filter(cohort == "TS cohort")) %>% 
ggplot(aes(group, ageAcc2, color = group, fill = group)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey80")+
  geom_point(alpha = .4, position = position_jitter(width = .35)) +
  geom_errorbar(aes(x = group, y = emmean, ymin = lower.CL, ymax = upper.CL), 
               stat = "identity", width = .35, color = "black") +
    geom_point(aes(x = group, y = emmean, fill = group),
             size = 2, alpha = 1, shape = 21, color = "black") + 
  annotate("text", x = 1.5, y = 15, label = paste("p =", signif(p.value, 2)), size = 3, 
            vjust = -0.5, color = "black") +
  theme_bw() + theme(legend.position = "none",
                     plot.title = element_text(hjust = .5))+
  scale_color_manual(values = col) + 
  scale_fill_manual(values = col) +
  labs(x = "", y = "Acceleration (years)", title = "ageAcc2") +
  ylim(-8,17) -> TSa2

emm <- emmeans(lm(ageAcc3 ~ group, data=acc_df_mean %>% filter(cohort == "TS cohort")), pairwise ~ group)
p.value <- as.data.frame(emm$contrasts)$p.value

merge(emm$emmeans,acc_df_mean %>% filter(cohort == "TS cohort")) %>% 
ggplot(aes(group, ageAcc3, color = group, fill = group)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey80")+
  geom_point(alpha = .4, position = position_jitter(width = .35)) +
  geom_errorbar(aes(x = group, y = emmean, ymin = lower.CL, ymax = upper.CL), 
               stat = "identity", width = .35, color = "black") +
    geom_point(aes(x = group, y = emmean, fill = group),
             size = 2, alpha = 1, shape = 21, color = "black") + 
  annotate("text", x = 1.5, y = 15, label = paste("p =", signif(p.value, 2)), size = 3, 
            vjust = -0.5, color = "black") +
  theme_bw() + theme(legend.position = "none",
                     plot.title = element_text(hjust = .5))+
  scale_color_manual(values = col) + 
  scale_fill_manual(values = col) +
  labs(x = "", y = "Acceleration (years)", title = "ageAcc3") +
  ylim(-8,17) -> TSa3

# KS
emm <- emmeans(lm(ageAcc ~ group, data=acc_df_mean %>% filter(cohort == "KS cohort")), pairwise ~ group)
p.value <- as.data.frame(emm$contrasts)$p.value

merge(emm$emmeans,acc_df_mean %>% filter(cohort == "KS cohort")) %>% 
ggplot(aes(group, ageAcc, color = group, fill = group)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey80")+
  geom_point(alpha = .4, position = position_jitter(width = .35)) +
  geom_errorbar(aes(x = group, y = emmean, ymin = lower.CL, ymax = upper.CL), 
               stat = "identity", width = .35, color = "black") +
    geom_point(aes(x = group, y = emmean, fill = group),
             size = 2, alpha = 1, shape = 21, color = "black") + 
  annotate("text", x = 1.5, y = 15, label = paste("p =", signif(p.value, 2)), size = 3, 
            vjust = -0.5, color = "black") +
  theme_bw() + theme(legend.position = "none",
                     plot.title = element_text(hjust = .5))+
  scale_color_manual(values = col) + 
  scale_fill_manual(values = col) +
  labs(x = "", y = "Acceleration (years)", title = "ageAcc") +
  ylim(-8,17) -> KSa1

emm <- emmeans(lm(ageAcc2 ~ group, data=acc_df_mean %>% filter(cohort == "KS cohort")), pairwise ~ group)
p.value <- as.data.frame(emm$contrasts)$p.value

merge(emm$emmeans,acc_df_mean %>% filter(cohort == "KS cohort")) %>% 
ggplot(aes(group, ageAcc2, color = group, fill = group)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey80")+
  geom_point(alpha = .4, position = position_jitter(width = .35)) +
  geom_errorbar(aes(x = group, y = emmean, ymin = lower.CL, ymax = upper.CL), 
               stat = "identity", width = .35, color = "black") +
    geom_point(aes(x = group, y = emmean, fill = group),
             size = 2, alpha = 1, shape = 21, color = "black") + 
  annotate("text", x = 1.5, y = 15, label = paste("p =", signif(p.value, 2)), size = 3, 
            vjust = -0.5, color = "black") +
  theme_bw() + theme(legend.position = "none",
                     plot.title = element_text(hjust = .5))+
  scale_color_manual(values = col) + 
  scale_fill_manual(values = col) +
  labs(x = "", y = "Acceleration (years)", title = "ageAcc2") +
  ylim(-8,17) -> KSa2

emm <- emmeans(lm(ageAcc3 ~ group, data=acc_df_mean %>% filter(cohort == "KS cohort")), pairwise ~ group)
p.value <- as.data.frame(emm$contrasts)$p.value

merge(emm$emmeans,acc_df_mean %>% filter(cohort == "KS cohort")) %>% 
ggplot(aes(group, ageAcc3, color = group, fill = group)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey80")+
  geom_point(alpha = .4, position = position_jitter(width = .35)) +
  geom_errorbar(aes(x = group, y = emmean, ymin = lower.CL, ymax = upper.CL), 
               stat = "identity", width = .35, color = "black") +
    geom_point(aes(x = group, y = emmean, fill = group),
             size = 2, alpha = 1, shape = 21, color = "black") + 
  annotate("text", x = 1.5, y = 15, label = paste("p =", signif(p.value, 2)), size = 3, 
            vjust = -0.5, color = "black") +
  theme_bw() + theme(legend.position = "none",
                     plot.title = element_text(hjust = .5))+
  scale_color_manual(values = col) + 
  scale_fill_manual(values = col) +
  labs(x = "", y = "Acceleration (years)", title = "ageAcc3") +
  ylim(-8,17) -> KSa3


```

```{r}

ggarrange(
  
  ggarrange(DNAmAge_mean_p_TS,
          DNAmAge_mean_p_TS2,
          DNAmAge_mean_p_KS,
          DNAmAge_mean_p_KS2,
          ncol = 4,
          labels = LETTERS[2:5],
          widths = c(1,1.5,1,1.5)),
  
  ggarrange(TSa1, 
          TSa2, 
          TSa3, 
          KSa1, 
          KSa2, 
          KSa3, 
          ncol = 6, labels = LETTERS[6:11]),
  
  ncol = 1, nrow = 2
) -> F1
F1

ggsave(F1, file = paste0(path_plots, "F1.pdf"),
                         width = 10, height = 6.25)

```

## Figure 2

```{r}
pheno_df <- df %>% 
  select(DNAmPhenoAge, AgeAccelPheno, cAge, group, cohort) %>% 
  rename("AgeAccelPheno" = "ageAccPheno")

grim_df <- df %>% 
  select(DNAmGrimAge2BasedOnRealAge, AgeAccelGrim, cAge, group, cohort) %>% 
  rename("DNAmGrimAge2BasedOnRealAge" = "DNAmGrimAge","AgeAccelGrim" = "ageAccGrim")

pace_df <- df %>% 
  select(DunedinPACE, cAge, group, cohort)

TL_df <- df %>% 
  select(TL, cAge, group, cohort)
```

### PhenoAge

```{r}
emm <- emmeans(lm(DNAmPhenoAge ~ cAge + group, data = pheno_df %>% filter(cohort == "TS cohort")), ~ group)
merge(
  pheno_df %>% filter(cohort == "TS cohort"),
  as.data.frame(emm),
  by = "group") %>% 
ggplot() +
  geom_jitter(aes(x = factor(group, levels = order), y = DNAmPhenoAge,fill = group, col = group), width = .4, alpha = .4) +
   geom_errorbar(aes(x = group, y = emmean, ymin = lower.CL, ymax = upper.CL), 
               stat = "identity", width = .35, color = "black") +
    geom_point(aes(x = group, y = emmean, fill = group),
             size = 2, alpha = .5, shape = 21, color = "black") + 
  annotate("text", x = 1.5, y = 65, label = paste("p =", signif(summary(contrast(emm, method = "pairwise"))$p.value, 2)), size = 3, vjust = -0.5) +
  theme_bw() + theme(legend.position = "none") +
  scale_color_manual(values = col) + 
  scale_fill_manual(values = col) +
  labs(x = "") +
  ylim(-5,70)-> phenoTS_p1

pheno_df %>% 
  filter(cohort == "TS cohort") %>% 
  ggplot(aes(cAge, DNAmPhenoAge, color = factor(group, levels = order))) +
  geom_point(alpha = .35) +
  geom_smooth(method = "lm", se = F) + 
  theme_bw()+theme(legend.position = "none") + 
  scale_color_manual(values = col) + 
  labs(x = "cAge")+
  ylim(-5,70) -> phenoTS_p2
```

```{r}
emm <- emmeans(lm(DNAmPhenoAge ~ cAge + group, data = pheno_df %>% filter(cohort == "KS cohort")), ~ group)
merge(
  pheno_df %>% filter(cohort == "KS cohort"),
  as.data.frame(emm),
  by = "group") %>% 
ggplot() +
  geom_jitter(aes(x = factor(group, levels = order), y = DNAmPhenoAge,fill = group, col = group), width = .4, alpha = .4) +
   geom_errorbar(aes(x = group, y = emmean, ymin = lower.CL, ymax = upper.CL), 
               stat = "identity", width = .35, color = "black") +
    geom_point(aes(x = group, y = emmean, fill = group),
             size = 2, alpha = .5, shape = 21, color = "black") + 
  annotate("text", x = 1.5, y = 65, label = paste("p =", signif(summary(contrast(emm, method = "pairwise"))$p.value, 2)), size = 3, vjust = -0.5) +
  theme_bw() + theme(legend.position = "none") +
  scale_color_manual(values = col) + 
  scale_fill_manual(values = col) +
  labs(x = "") +
  ylim(-5,70)-> phenoKS_p1

pheno_df %>% 
  filter(cohort == "KS cohort") %>% 
  ggplot(aes(cAge, DNAmPhenoAge, color = factor(group, levels = order))) +
  geom_point(alpha = .35) +
  geom_smooth(method = "lm", se = F) + 
  theme_bw()+theme(legend.position = "none") + 
  scale_color_manual(values = col) + 
  labs(x = "cAge") +
  ylim(-5,70)-> phenoKS_p2
```

```{r}

emm <- emmeans(lm(ageAccPheno ~ group, data = pheno_df %>% filter(cohort == "TS cohort")), ~ group)
merge(
  pheno_df %>% filter(cohort == "TS cohort"),
  as.data.frame(emm),
  by = "group") %>% 
ggplot() +
    geom_hline(yintercept = 0, linetype = "dashed", color = "grey80") + 
  geom_jitter(aes(x = factor(group, levels = order), y = ageAccPheno,fill = group, col = group), width = .4, alpha = .4) +
   geom_errorbar(aes(x = group, y = emmean, ymin = lower.CL, ymax = upper.CL), 
               stat = "identity", width = .35, color = "black") +
    geom_point(aes(x = group, y = emmean, fill = group),
             size = 2, alpha = .5, shape = 21, color = "black") + 
  annotate("text", x = 1.5, y = 20, label = paste("p =", signif(summary(contrast(emm, method = "pairwise"))$p.value, 2)), size = 3, vjust = -0.5) +
  theme_bw() + theme(legend.position = "none") +
  scale_color_manual(values = col) + 
  scale_fill_manual(values = col) +
  labs(x = "") +
  ylim(-32,25)-> phenoTS_p3

emm <- emmeans(lm(ageAccPheno ~ group, data = pheno_df %>% filter(cohort == "KS cohort")), ~ group)
merge(
  pheno_df %>% filter(cohort == "KS cohort"),
  as.data.frame(emm),
  by = "group") %>% 
ggplot() +
    geom_hline(yintercept = 0, linetype = "dashed", color = "grey80") + 
  geom_jitter(aes(x = factor(group, levels = order), y = ageAccPheno,fill = group, col = group), width = .4, alpha = .4) +
   geom_errorbar(aes(x = group, y = emmean, ymin = lower.CL, ymax = upper.CL), 
               stat = "identity", width = .35, color = "black") +
    geom_point(aes(x = group, y = emmean, fill = group),
             size = 2, alpha = .5, shape = 21, color = "black") + 
  annotate("text", x = 1.5, y = 20, label = paste("p =", signif(summary(contrast(emm, method = "pairwise"))$p.value, 2)), size = 3, vjust = -0.5) +
  theme_bw() + theme(legend.position = "none") +
  scale_color_manual(values = col) + 
  scale_fill_manual(values = col) +
  labs(x = "") +
  ylim(-32,25) -> phenoKS_p3
```

### GrimAge

```{r}
emm <- emmeans(lm(DNAmGrimAge ~ cAge + group, data = grim_df %>% filter(cohort == "TS cohort")), ~ group)
merge(
  grim_df %>% filter(cohort == "TS cohort"),
  as.data.frame(emm),
  by = "group") %>% 
ggplot() +
  geom_jitter(aes(x = factor(group, levels = order), y = DNAmGrimAge,fill = group, col = group), width = .4, alpha = .4) +
   geom_errorbar(aes(x = group, y = emmean, ymin = lower.CL, ymax = upper.CL), 
               stat = "identity", width = .35, color = "black") +
    geom_point(aes(x = group, y = emmean, fill = group),
             size = 2, alpha = .5, shape = 21, color = "black") + 
  annotate("text", x = 1.5, y = 80, label = paste("p =", signif(summary(contrast(emm, method = "pairwise"))$p.value, 2)), size = 3, vjust = -0.5) +
  theme_bw() + theme(legend.position = "none") +
  scale_color_manual(values = col) + 
  scale_fill_manual(values = col) +
  labs(x = "") +
  ylim(25,85)-> grimTS_p1

grim_df %>% 
  filter(cohort == "TS cohort") %>% 
  ggplot(aes(cAge, DNAmGrimAge, color = factor(group, levels = order))) +
  geom_point(alpha = .35) +
  geom_smooth(method = "lm", se = F) + 
  theme_bw()+theme(legend.position = "none") + 
  scale_color_manual(values = col) + 
  labs(x = "cAge")+
  ylim(25,85) -> grimTS_p2
```

```{r}
emm <- emmeans(lm(DNAmGrimAge ~ cAge + group, data = grim_df %>% filter(cohort == "KS cohort")), ~ group)
merge(
  grim_df %>% filter(cohort == "KS cohort"),
  as.data.frame(emm),
  by = "group") %>% 
ggplot() +
  geom_jitter(aes(x = factor(group, levels = order), y = DNAmGrimAge,fill = group, col = group), width = .4, alpha = .4) +
   geom_errorbar(aes(x = group, y = emmean, ymin = lower.CL, ymax = upper.CL), 
               stat = "identity", width = .35, color = "black") +
    geom_point(aes(x = group, y = emmean, fill = group),
             size = 2, alpha = .5, shape = 21, color = "black") + 
  annotate("text", x = 1.5, y = 80, label = paste("p =", signif(summary(contrast(emm, method = "pairwise"))$p.value, 2)), size = 3, vjust = -0.5) +
  theme_bw() + theme(legend.position = "none") +
  scale_color_manual(values = col) + 
  scale_fill_manual(values = col) +
  labs(x = "") +
  ylim(25,85)-> grimKS_p1

grim_df %>% 
  filter(cohort == "KS cohort") %>% 
  ggplot(aes(cAge, DNAmGrimAge, color = factor(group, levels = order))) +
  geom_point(alpha = .35) +
  geom_smooth(method = "lm", se = F) + 
  theme_bw()+theme(legend.position = "none") + 
  scale_color_manual(values = col) + 
  labs(x = "cAge") +
  ylim(25,85)-> grimKS_p2
```

```{r}

emm <- emmeans(lm(ageAccGrim ~ group, data = grim_df %>% filter(cohort == "TS cohort")), ~ group)
merge(
  grim_df %>% filter(cohort == "TS cohort"),
  as.data.frame(emm),
  by = "group") %>% 
ggplot() +
    geom_hline(yintercept = 0, linetype = "dashed", color = "grey80") + 
  geom_jitter(aes(x = factor(group, levels = order), y = ageAccGrim,fill = group, col = group), width = .4, alpha = .4) +
   geom_errorbar(aes(x = group, y = emmean, ymin = lower.CL, ymax = upper.CL), 
               stat = "identity", width = .35, color = "black") +
    geom_point(aes(x = group, y = emmean, fill = group),
             size = 2, alpha = .5, shape = 21, color = "black") + 
  annotate("text", x = 1.5, y = 15, label = paste("p =", signif(summary(contrast(emm, method = "pairwise"))$p.value, 2)), size = 3, vjust = -0.5) +
  theme_bw() + theme(legend.position = "none") +
  scale_color_manual(values = col) + 
  scale_fill_manual(values = col) +
  labs(x = "")  +
  ylim(-10,16)-> grimTS_p3

emm <- emmeans(lm(ageAccGrim ~ group, data = grim_df %>% filter(cohort == "KS cohort")), ~ group)
merge(
  grim_df %>% filter(cohort == "KS cohort"),
  as.data.frame(emm),
  by = "group") %>% 
ggplot() +
    geom_hline(yintercept = 0, linetype = "dashed", color = "grey80") + 
  geom_jitter(aes(x = factor(group, levels = order), y = ageAccGrim,fill = group, col = group), width = .4, alpha = .4) +
   geom_errorbar(aes(x = group, y = emmean, ymin = lower.CL, ymax = upper.CL), 
               stat = "identity", width = .35, color = "black") +
    geom_point(aes(x = group, y = emmean, fill = group),
             size = 2, alpha = .5, shape = 21, color = "black") + 
  annotate("text", x = 1.5, y = 15, label = paste("p =", signif(summary(contrast(emm, method = "pairwise"))$p.value, 2)), size = 3, vjust = -0.5) +
  theme_bw() + theme(legend.position = "none") +
  scale_color_manual(values = col) + 
  scale_fill_manual(values = col) +
  labs(x = "") +
  ylim(-10,16)-> grimKS_p3
```

```{r}

ggarrange(
  
  ggarrange(phenoTS_p1,phenoTS_p2,
            phenoKS_p1,phenoKS_p2,
          ncol = 4, nrow = 1, widths = c(1,1.5,1,1.5),
          labels = c("A","C","E","G")),
  
  ggarrange(grimTS_p1,grimTS_p2,
            grimKS_p1,grimKS_p2,
          ncol = 4, nrow = 1, widths = c(1,1.5,1,1.5),
          labels = c("B","D","F","H")),
  
  ggarrange(phenoTS_p3,phenoKS_p3,
          grimTS_p3,grimKS_p3,NA,
          ncol = 5, nrow = 1,
          labels = c("I","J","K","L",NA)),
  
  ncol = 1, nrow = 3
) -> F2
F2

ggsave(F2, file = paste0(path_plots, "F2.pdf"),
       width = 10, height = 9.375)

```

## Figure 3

### DNAmTL

```{r TL KS 450K, message=FALSE, warning=FALSE, fig.height=3, fig.width=3}

# Make model for each cohort and calculate EM-means:
emm_TL_TS <- emmeans(lm(TL ~ cAge + group,  data = TL_df %>% filter(cohort == "TS cohort") ),~ group)
merge(TL_df,
      as.data.frame(emm_TL_TS) %>%                  # Make combined data frame and add p-values
        mutate(p_value = summary(contrast(emm_TL_TS, method = "pairwise"))$p.value),
      by = "group") %>% 
  ggplot() +
  geom_jitter(aes(x = factor(group, levels = order), y = TL,fill = group, col = group), 
              width = .35, alpha = .4) +
  geom_errorbar(aes(x = group, y = emmean, ymin = lower.CL, ymax = upper.CL), 
               stat = "identity", width = .35, color = "black") +
    geom_point(aes(x = group, y = emmean, fill = group),
             size = 2, alpha = 1, shape = 21, color = "black") + 
  annotate("text", x = 1.5, y = 8,label = paste("p =", signif(summary(contrast(emm_TL_TS, method = "pairwise"))$p.value, 2)), size = 3, vjust = -0.5) +
  theme_bw() + theme(legend.position = "none") +
  scale_color_manual(values = col) + 
  scale_fill_manual(values = col) +
  ylim(6.25,8.25) +
  labs(x = "", y = "Telomere Length (kb)") -> TL_p_TS

TL_df %>% 
  filter(cohort == "TS cohort") %>% 
  ggplot(aes(cAge, TL, color = factor(group, levels = order))) +
  geom_point(alpha = .35) +
  geom_smooth(method = "lm", se = F) + 
  theme_bw()+theme(legend.position = "none") + 
  scale_color_manual(values = col) + 
  ylim(6.25,8.25) +
  labs(x = "cAge", y = "Telomere Length (kb)") -> TL_p_lm_TS

```

```{r}
emm_TL_KS <- emmeans(lm(TL ~ cAge + group, data = TL_df %>% filter(cohort == "KS cohort") ), ~ group)
merge(TL_df,
      as.data.frame(emm_TL_KS) %>%                  # Make combined data frame and add p-values
        mutate(p_value = summary(contrast(emm_TL_KS, method = "pairwise"))$p.value),
      by = "group") %>% 
  ggplot() +
  geom_jitter(aes(x = factor(group, levels = order), y = TL, fill = group, col = group), 
              width = .35, alpha = .4) +
  geom_errorbar(aes(x = group, y = emmean, ymin = lower.CL, ymax = upper.CL), 
               stat = "identity", width = .35, color = "black") +
    geom_point(aes(x = group, y = emmean, fill = group),
             size = 2, alpha = 1, shape = 21, color = "black") + 
  annotate("text", x = 1.5, y = 8,label = paste("p =", signif(summary(contrast(emm_TL_KS, method = "pairwise"))$p.value, 2)), size = 3, vjust = -0.5) +
  theme_bw() + theme(legend.position = "none") +
  scale_color_manual(values = col) + 
  scale_fill_manual(values = col) +
  ylim(6.25,8.25) +
  labs(x = "", y = "Telomere Length (kb)")  -> TL_p_KS


TL_df %>% 
  filter(cohort == "KS cohort") %>% 
  ggplot(aes(cAge, TL, color = factor(group, levels = order))) +
  geom_point(alpha = .35) +
  geom_smooth(method = "lm", se = F) + 
  theme_bw()+theme(legend.position = "none") + 
  scale_color_manual(values = col) + 
  ylim(6.25,8.25) +
  labs(x = "cAge", y = "Telomere Length (kb)") -> TL_p_lm_KS
```

### DunedinPACE

```{r message=FALSE, warning=FALSE, fig.height=4, fig.width=4}

emm_pace_TS <- emmeans(lm(DunedinPACE ~ group, data = pace_df %>% filter(cohort == "TS cohort") ), ~ group)
merge(pace_df,as.data.frame(emm_pace_TS),by = "group") %>% 
  ggplot() +
  geom_hline(yintercept = 1, color = "grey80", linetype = "dashed") +
  geom_jitter(aes(x = factor(group, levels = order), y = DunedinPACE, fill = group, col = group), 
              width = .35, alpha = .4) +
  geom_errorbar(aes(x = group, y = emmean, ymin = lower.CL, ymax = upper.CL), 
               stat = "identity", width = .35, color = "black") +
    geom_point(aes(x = group, y = emmean, fill = group),
             size = 2, alpha = 1, shape = 21, color = "black") + 
  annotate("text", x = 1.5, y = 1.4,label = paste("p =", signif(summary(contrast(emm_pace_TS, method = "pairwise"))$p.value, 2)), size = 3, vjust = -0.5) +
  theme_bw() + theme(legend.position = "none") +
  scale_color_manual(values = col) + 
  scale_fill_manual(values = col) +
  labs(x = "", y = "DunedinPACE") +
  ylim(0.5,1.5) -> PACE_p_TS

emm_pace_KS <- emmeans(lm(DunedinPACE ~ group, data = pace_df %>% filter(cohort == "KS cohort") ), ~ group)
merge(pace_df,as.data.frame(emm_pace_KS),by = "group") %>% 
  ggplot() +
  geom_hline(yintercept = 1, color = "grey80", linetype = "dashed") +
  geom_jitter(aes(x = factor(group, levels = order), y = DunedinPACE, fill = group, col = group), 
              width = .35, alpha = .4) +
  geom_errorbar(aes(x = group, y = emmean, ymin = lower.CL, ymax = upper.CL), 
               stat = "identity", width = .35, color = "black") +
    geom_point(aes(x = group, y = emmean, fill = group),
             size = 2, alpha = 1, shape = 21, color = "black") + 
  annotate("text", x = 1.5, y = 1.4,label = paste("p =", signif(summary(contrast(emm_pace_KS, method = "pairwise"))$p.value, 2)), size = 3, vjust = -0.5) +
  theme_bw() + theme(legend.position = "none") +
  scale_color_manual(values = col) + 
  scale_fill_manual(values = col) +
  labs(x = "", y = "DunedinPACE") +
  ylim(0.5,1.5) -> PACE_p_KS


pace_df %>%  
  filter(cohort == "TS cohort") %>% 
  ggplot(aes(cAge, DunedinPACE, color = group, fill = group)) +
  geom_hline(yintercept = 1, color = "grey80", linetype = "dashed") +
  geom_point(alpha = .4) +
  geom_smooth(method = "lm", se = F, linewidth = .75) +
  theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5),
                     legend.position = "none") +
  labs(x = "cAge", y = "DunedinPACE") + 
  ylim(0.5,1.5)+
  scale_color_manual(values = col) + 
  scale_fill_manual(values = col) -> PACE_p_lm_TS

pace_df %>%  
  filter(cohort == "KS cohort") %>% 
  ggplot(aes(cAge, DunedinPACE, color = group, fill = group)) +
  geom_hline(yintercept = 1, color = "grey80", linetype = "dashed") +
  geom_point(alpha = .4) +
  geom_smooth(method = "lm", se = F, linewidth = .75) +
  theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5),
                     legend.position = "none") +
  labs(x = "cAge", y = "DunedinPACE") + 
  ylim(0.5,1.5)+
  scale_color_manual(values = col) + 
  scale_fill_manual(values = col) -> PACE_p_lm_KS
```

## Clin

```{r}

CLIN_TS <- readRDS("/path/clinical_traits.rds") %>% mutate("Waist:Hip-ratio" = (`Waist (cm)`/`Hip (cm)`))
CLIN_KS <- readRDS("/path/KS_clinical.rds")

df_mean <- df %>% 
  mutate(DNAmAge = rowMeans(select(df, BLUP, EN, Hannum, Horvath)),
         ageAcc = rowMeans(select(df, ageAcc.BLUP, ageAcc.EN, ageAcc.Hannum, ageAcc.Horvath)),
         ageAcc2 = rowMeans(select(df, ageAcc2.BLUP, ageAcc2.EN, ageAcc2.Hannum, ageAcc2.Horvath)),
         ageAcc3 = rowMeans(select(df, ageAcc3.BLUP, ageAcc3.EN, ageAcc3.Hannum, ageAcc3.Horvath))) 

vars2 = c("cAge",
         "DNAmAge",
         "PhenoAge","GrimAge",
         "DNAmTL","DunedinPACE", 
         "ageAcc",
         "ageAcc2",
         "ageAcc3",
        "ageAccPheno","ageAccGrim")

traits_df_TS = data.frame(
  traits = c(# Body composition
           "Weight (kg)",
           "BMI (kg/m2)",
           "Hip (cm)",
           "Waist (cm)",
           "Waist-Hip-ratio",
           
           #Cardiovasculature
           "Systolic 24h (mmHg)",
           "Diastolic 24h (mmHg)",
           "TotalPlaque",
           "ECG-QTc (ms)",
           "Left ventricular mass (g)",
           "Ejection fraction %",
           "E/A ratio",
           
           # Lipids
           "HDL cholesterol (mmol/l)",
           "Triglycerides (mmol/l)",
           "Total cholesterol (mmol/l)",
           
           # Biochemistry
           "Hemoglobin (mmol/l)",
           "Creatinine (µmol/l)",
           "HbA1c (mmol/mol)"),
  
  trait_groups = c(rep("Body composition",5),
                 rep("Cardiovasculature", 7),
                 rep("Lipids",3),
                 rep("Biochemistry",3)))
  
```

```{r}
merge(
  merge(
    df_mean %>% 
      rename(TL = "DNAmTL",
             DNAmGrimAge2BasedOnRealAge = "GrimAge",
             AgeAccelGrim = "ageAccGrim",
             DNAmPhenoAge = "PhenoAge",
             AgeAccelPheno = "ageAccPheno") %>% 
      filter(cohort == "TS cohort") %>% 
      select(vars2, id, group, cAge) %>% 
      pivot_longer(-c(id, group), names_to = "Algorithm", values_to = "value"),
    TS450K_info %>% select(ID_new, somaseq) %>% 
            mutate(ID_new = gsub("-",".", ID_new),
                   ID_new = gsub(" ",".", ID_new)),
    by.x = "id", by.y = "ID_new"),
  CLIN,
  by.x = "somaseq", by.y = "somaseq_id") %>% 
  select(group.x, Algorithm, value, traits_df$traits) -> CLIN_df_M_mean

```

```{r}
cor_df_TS_m <- CLIN_df_M_mean %>% 
  filter(group.x == "TS") %>% 
  pivot_wider(names_from = Algorithm, values_from = value) %>% 
  select(-group.x) %>%
  as.data.frame() %>% 
  as.matrix() %>%
  Hmisc::rcorr()

r_TS_m <- cor_df_TS_m[[1]] %>% 
    as.data.frame() %>% 
    filter(row.names(.) %in% vars2) %>% 
    mutate(algorithms = rownames(.)) %>% 
    pivot_longer(-algorithms)

p_TS_m <- cor_df_TS_m[[3]] %>% 
  as.data.frame() %>% 
  filter(row.names(.) %in% vars2) %>% 
    mutate(algorithms = rownames(.)) %>% 
    pivot_longer(-algorithms)

order_cor_TS_m <- unique((r_TS_m %>%  
                  filter(algorithms == "cAge") %>% 
                  arrange(value))$name )

cor_df_F_m <- CLIN_df_M_mean %>% 
    filter(group.x == "Female") %>% 
  pivot_wider(names_from = Algorithm, values_from = value) %>% 
  select(-group.x) %>%
  as.data.frame() %>% 
  as.matrix() %>%
  Hmisc::rcorr()

r_F_m <- cor_df_F_m[[1]] %>% 
    as.data.frame() %>% 
    filter(row.names(.) %in% vars2) %>% 
    mutate(algorithms = rownames(.)) %>% 
    pivot_longer(-algorithms)

p_F_m <- cor_df_F_m[[3]] %>% as.data.frame() %>% 
  filter(row.names(.) %in% vars2) %>% 
    mutate(algorithms = rownames(.)) %>% 
    pivot_longer(-algorithms)

cor_r_p_m <- rbind(
  r_TS_m %>% mutate(p = p_TS_m$value,
                  p_ = case_when(p < 0.05 ~ "*" , .default = "")) %>% 
    mutate(group = "TS"),
  
  r_F_m %>% mutate(p = p_F_m$value,
                 p_ = case_when(p < 0.05 ~ "*" , .default = "")) %>% 
    mutate(group = "Female")) 
  
merge(cor_r_p_m,
      traits_df,
      by.x = "name",
      by.y = "traits") %>% 
ggplot(aes(factor(name, levels = traits_df$traits), 
           factor(algorithms, levels = vars2), 
           fill = value)) + 
  geom_tile() + 
  geom_text(aes(label = paste(round(value, digits = 2),
                              "\n",
                              ifelse(p < 0.001, "<0.001",round(p, digits = 3))), color = p_), size = 1.5) + 
  scale_color_manual(values = c("grey","black")) + 
  facet_grid(rows = vars(factor(group, levels = c("TS", "Female"))),
             cols = vars(trait_groups),
             scales = "free",
             space = "free") +
  theme_bw() +
  scale_fill_gradient2(mid = "white", high = col[1], low = col[2]) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5),
        legend.position = "right") +
  labs(x = "", y = "", title = "") -> cor_p_TS_mean
cor_p_TS_mean
```

```{r}

traits_KS_df = data.frame(
  traits = c(# Body composition
           "Weight (kg)",
           "BMI (kg/m2)",
           "Hip (cm)",
           "Waist (cm)",
           "Waist-Hip-Ratio",
           "Body Fat %", 
           
           # Lipids
           "HDL cholesterol (mmol/l)",
           "Triglycerides (mmol/l)",
           "Total cholesterol (mmol/l)"),
  
  trait_groups = c(rep("Body composition",6),
                 rep("Lipids",3)))
merge(
  merge(
    df_mean %>% 
      rename(TL = "DNAmTL",
        DNAmGrimAge2BasedOnRealAge = "GrimAge",
        AgeAccelGrim = "ageAccGrim",
        DNAmPhenoAge = "PhenoAge",
        AgeAccelPheno = "ageAccPheno") %>% 
      filter(cohort == "KS cohort") %>% 
      select(vars2, id, group, cAge) %>% 
      pivot_longer(-c(id, group), names_to = "Algorithm", values_to = "value"),
  KS450K_info %>% select(unique_id, slide_array),
  by.x = "id", by.y = "slide_array"),
  CLIN_KS, by = "unique_id") %>% 
  select(-c(Diagnosis_Age, Treatment_start_age, Testis_vol)) %>% 
  rename("Weight" = "Weight (kg)",
          "BMI"="BMI (kg/m2)",
          "Hip"="Hip (cm)" ,
          "Waist"="Waist (cm)",
          "Waist_Hip_Ratio" ="Waist-Hip-Ratio",
          "HDL" ="HDL cholesterol (mmol/l)",
          "Triglycerides" ="Triglycerides (mmol/l)",
          "Total_cholesterol"= "Total cholesterol (mmol/l)",
          "hemoglobin"="Hemoglobin (mmol/l)",
         "Body_fat_percentage" = "Body Fat %")  -> CLIN_df_M_KS_mean
```

```{r}
cor_df_KS <- CLIN_df_M_KS_mean %>% 
  filter(group == "KS") %>% 
  select(-c(id, group, unique_id, Group, Karyotype, Treatment, Age)) %>%
  pivot_wider(names_from = Algorithm, values_from = value) %>% 
  as.data.frame() %>% 
  as.matrix() %>%
  Hmisc::rcorr()

r_KS <- cor_df_KS[[1]] %>% 
    as.data.frame() %>% 
    filter(row.names(.) %in% c(vars2, "cAge")) %>% 
    select(traits_KS_df$traits) %>% 
    mutate(algorithms = rownames(.)) %>% 
    pivot_longer(-algorithms)

p_KS <- cor_df_KS[[3]] %>% 
    as.data.frame() %>% 
    filter(row.names(.) %in% c(vars2, "cAge")) %>% 
  select(traits_KS_df$traits) %>% 
    mutate(algorithms = rownames(.)) %>% 
    pivot_longer(-algorithms)

order_cor_KS <- unique((r_KS %>%  
                  filter(algorithms == "cAge") %>% 
                  arrange(value))$name )

cor_df_M <- CLIN_df_M_KS_mean %>% 
  filter(group == "Male") %>% 
  select(-c(id, group, unique_id, Group, Karyotype, Treatment, Age)) %>%
  pivot_wider(names_from = Algorithm, values_from = value) %>% 
  as.data.frame() %>% 
  as.matrix() %>%
  Hmisc::rcorr()

r_M <- cor_df_M[[1]] %>% 
    as.data.frame() %>% 
    filter(row.names(.) %in% c(vars2, "cAge")) %>% 
  select(traits_KS_df$traits) %>% 
    mutate(algorithms = rownames(.)) %>% 
    pivot_longer(-algorithms)

p_M <- cor_df_M[[3]] %>% 
  as.data.frame() %>% 
    filter(row.names(.) %in% c(vars2, "cAge")) %>% 
  select(traits_KS_df$traits) %>% 
    mutate(algorithms = rownames(.)) %>% 
    pivot_longer(-algorithms)

cor_r_p_KScohort <- merge(
  rbind(
  r_KS %>% mutate(p = p_KS$value,
                  p_ = case_when(p < 0.05 ~ "*" , .default = "")) %>% 
    mutate(group = "KS"),
  
  r_M %>% mutate(p = p_M$value,
                 p_ = case_when(p < 0.05 ~ "*" , .default = "")) %>% 
    mutate(group = "Male")),
  traits_KS_df,
  by.x = "name", by.y = "traits")
  
cor_r_p_KScohort %>% 
ggplot(aes(factor(name, levels = traits_KS_df$traits),
           factor(algorithms, levels = c(vars2)),
           fill = value)) + 
  geom_tile() + 
  geom_text(aes(label = paste(round(value, digits = 2),
                            "\n",
                            ifelse(p < 0.001, "<0.001",round(p, digits = 3))), color = p_), size = 1.5) + 
  scale_color_manual(values = c("grey","black")) + 
  facet_grid(rows = vars(factor(group, levels = c("KS","Male"))),
             cols = vars(trait_groups),
             scales = "free",
             space = "free") +
  theme_bw() +
  scale_fill_gradient2(mid = "white", high = col[4], low = col[3]) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5),
        legend.position = "right") +
  labs(x = "", y = "", title = "") -> cor_p_KS_mean
cor_p_KS_mean

```

```{r}

ggarrange(
  
   ggarrange(PACE_p_TS, PACE_p_lm_TS, 
          PACE_p_KS, PACE_p_lm_KS,
          TL_p_TS, TL_p_lm_TS, 
          TL_p_KS,TL_p_lm_KS,
          ncol = 4, nrow = 2, widths = c(1,1.5,1,1.5),
          labels = LETTERS[1:8]),
   
   ggarrange(cor_p_TS_mean,
             cor_p_KS_mean,
             widths = c(1.75,1.25),
             labels = LETTERS[9:10],
             ncol = 2, nrow = 1),
   
   ncol = 1, nrow = 2
) -> F3
F3

ggsave(F3, file = paste0(path_plots, "F3.pdf"),
       width = 10,
       height = 12.5)
```

## Figure S1

```{r}
vars = c("cAge",
         "BLUP","EN","Hannum","Horvath",
         "PhenoAge","GrimAge",
         "DNAmTL","DunedinPACE", 
         "ageAcc.BLUP","ageAcc2.BLUP","ageAcc3.BLUP",
         "ageAcc.EN","ageAcc2.EN","ageAcc3.EN",
         "ageAcc.Hannum","ageAcc2.Hannum","ageAcc3.Hannum",
         "ageAcc.Horvath","ageAcc2.Horvath","ageAcc3.Horvath",
        "ageAccPheno","ageAccGrim")

merge(
  merge(
    df %>% 
      rename(TL = "DNAmTL",
             DNAmGrimAge2BasedOnRealAge = "GrimAge",
             AgeAccelGrim = "ageAccGrim",
             DNAmPhenoAge = "PhenoAge",
             AgeAccelPheno = "ageAccPheno") %>% 
      filter(cohort == "TS cohort") %>% 
      select(vars, id, group, cAge) %>% 
      pivot_longer(-c(id, group), names_to = "Algorithm", values_to = "value"),
    TS450K_info %>% select(ID_new, somaseq) %>% 
            mutate(ID_new = gsub("-",".", ID_new),
                   ID_new = gsub(" ",".", ID_new)),
    by.x = "id", by.y = "ID_new"),
  CLIN_TS,
  by.x = "somaseq", by.y = "somaseq_id") %>% 
  rename("Waist:Hip-ratio" = "Waist-Hip-ratio")%>% 
  select(group.x, Algorithm, value, traits_df_TS$traits, id)  -> CLIN_df_M_TS

# We construct a correlation matrix for the TS group:

cor_df_TS_m <- CLIN_df_M %>% 
  filter(!(Algorithm %in% c("Body Mass Index", "Body Fat %", "HDL Cholesterol", "Waist:Hip Ratio")),
         group.x == "TS") %>% 
  pivot_wider(names_from = Algorithm, values_from = value) %>% 
  select(-group.x) %>%
  as.data.frame() %>% 
  as.matrix() %>%
  Hmisc::rcorr()

r_TS_m <- cor_df_TS_m[[1]] %>% 
    as.data.frame() %>% 
    filter(row.names(.) %in% vars) %>% 
    mutate(algorithms = rownames(.)) %>% 
    pivot_longer(-algorithms)

p_TS_m <- cor_df_TS_m[[3]] %>% 
  as.data.frame() %>% 
  filter(row.names(.) %in% vars) %>% 
    mutate(algorithms = rownames(.)) %>% 
    pivot_longer(-algorithms)

order_cor_TS_m <- unique((r_TS_m %>%  
                  filter(algorithms == "cAge") %>% 
                  arrange(value))$name )

cor_df_F_m <- CLIN_df_M %>% 
    filter(!(Algorithm %in% c("Body Mass Index", "Body Fat %", "HDL Cholesterol", "Waist:Hip Ratio")),
           group.x == "Female") %>% 
  pivot_wider(names_from = Algorithm, values_from = value) %>% 
  select(-group.x) %>%
  as.data.frame() %>% 
  as.matrix() %>%
  Hmisc::rcorr()

r_F_m <- cor_df_F_m[[1]] %>% 
    as.data.frame() %>% 
    filter(row.names(.) %in% vars) %>% 
    mutate(algorithms = rownames(.)) %>% 
    pivot_longer(-algorithms)

p_F_m <- cor_df_F_m[[3]] %>% as.data.frame() %>% 
  filter(row.names(.) %in% vars) %>% 
    mutate(algorithms = rownames(.)) %>% 
    pivot_longer(-algorithms)

cor_r_p_m <- rbind(
  r_TS_m %>% mutate(p = p_TS_m$value,
                  p_ = case_when(p < 0.05 ~ "*" , .default = "")) %>% 
    mutate(group = "TS"),
  
  r_F_m %>% mutate(p = p_F_m$value,
                 p_ = case_when(p < 0.05 ~ "*" , .default = "")) %>% 
    mutate(group = "Female")) 
  
merge(cor_r_p_m,
      traits_df,
      by.x = "name",
      by.y = "traits") %>% 
ggplot(aes(factor(name, levels = order_cor_TS_m), 
           factor(algorithms, levels = vars), 
           fill = value)) + 
  geom_tile() + 
  geom_text(aes(label = ifelse(p < 0.001, "<0.001",round(p, digits = 3)), color = p_), size = 1.5) + 
  scale_color_manual(values = c("grey","black")) + 
  facet_grid(rows = vars(factor(group, levels = c("TS", "Female"))),
             cols = vars(trait_groups),
             scales = "free",
             space = "free") +
  theme_bw() +
  scale_fill_gradient2(mid = "white", high = col[1], low = col[2]) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5),
        legend.position = "right") +
  labs(x = "", y = "", title = "") -> cor_p_TS
cor_p_TS
```

```{r}
merge(
  merge(
    df %>% 
      rename(TL = "DNAmTL",
        DNAmGrimAge2BasedOnRealAge = "GrimAge",
        AgeAccelGrim = "ageAccGrim",
        DNAmPhenoAge = "PhenoAge",
        AgeAccelPheno = "ageAccPheno") %>% 
      filter(cohort == "KS cohort") %>% 
      select(vars, id, group, cAge, "Body Mass Index", "Body Fat %", "HDL Cholesterol", "Waist:Hip Ratio") %>% 
      pivot_longer(-c(id, group), names_to = "Algorithm", values_to = "value"),
  KS450K_info %>% select(unique_id, slide_array),
  by.x = "id", by.y = "slide_array"),
  CLIN_KS, by = "unique_id") %>% 
  select(-c(Diagnosis_Age, Treatment_start_age, Testis_vol)) %>% 
  rename("Weight" = "Weight (kg)",
          "BMI"="BMI (kg/m2)",
          "Hip"="Hip (cm)" ,
          "Waist"="Waist (cm)",
          "Waist_Hip_Ratio" ="Waist-Hip-Ratio",
          "HDL" ="HDL cholesterol (mmol/l)",
          "Triglycerides" ="Triglycerides (mmol/l)",
          "Total_cholesterol"= "Total cholesterol (mmol/l)",
          "hemoglobin"="Hemoglobin (mmol/l)") -> CLIN_df_M_KS

traits_KS_df = data.frame(
  traits = c(# Body composition
           "Weight (kg)",
           "BMI (kg/m2)",
           "Hip (cm)",
           "Waist (cm)",
           "Waist-Hip-Ratio",
           "Body_fat_percentage",
           
           # Triglycerides
           "HDL cholesterol (mmol/l)",
           "Triglycerides (mmol/l)",
           "Total cholesterol (mmol/l)",
           
           # Biochemistry
           "Hemoglobin (mmol/l)"),
  
  trait_groups = c(rep("Body composition",6),
                 rep("Triglycerides",3),
                 rep("Biochemistry",1)))
```

We construct a correlation matrix for the KS group:

```{r}
cor_df_KS <- CLIN_df_M_KS %>% 
  filter(!(Algorithm %in% c("Body Mass Index", "Body Fat %", "HDL Cholesterol", "Waist:Hip Ratio")),
           group == "KS") %>% 
  select(-c(id, group, unique_id, Group, Karyotype, Treatment, Age)) %>%
  pivot_wider(names_from = Algorithm, values_from = value) %>% 
  as.data.frame() %>% 
  as.matrix() %>%
  Hmisc::rcorr()

r_KS <- cor_df_KS[[1]] %>% 
    as.data.frame() %>% 
    filter(row.names(.) %in% c(vars, "cAge")) %>% 
    select(traits_KS_df$traits) %>% 
    mutate(algorithms = rownames(.)) %>% 
    pivot_longer(-algorithms)

p_KS <- cor_df_KS[[3]] %>% 
    as.data.frame() %>% 
    filter(row.names(.) %in% c(vars, "cAge")) %>% 
  select(traits_KS_df$traits) %>% 
    mutate(algorithms = rownames(.)) %>% 
    pivot_longer(-algorithms)

order_cor_KS <- unique((r_KS %>%  
                  filter(algorithms == "cAge") %>% 
                  arrange(value))$name )
```

We construct a correlation matrix for the male group:

```{r}
cor_df_M <- CLIN_df_M_KS %>% 
  filter(!(Algorithm %in% c("Body Mass Index", "Body Fat %", "HDL Cholesterol", "Waist:Hip Ratio")),
           group == "Male") %>% 
  select(-c(id, group, unique_id, Group, Karyotype, Treatment, Age)) %>%
  pivot_wider(names_from = Algorithm, values_from = value) %>% 
  as.data.frame() %>% 
  as.matrix() %>%
  Hmisc::rcorr()

r_M <- cor_df_M[[1]] %>% 
    as.data.frame() %>% 
    filter(row.names(.) %in% c(vars, "cAge")) %>% 
  select(traits_KS_df$traits) %>% 
    mutate(algorithms = rownames(.)) %>% 
    pivot_longer(-algorithms)

p_M <- cor_df_M[[3]] %>% 
  as.data.frame() %>% 
    filter(row.names(.) %in% c(vars, "cAge")) %>% 
  select(traits_KS_df$traits) %>% 
    mutate(algorithms = rownames(.)) %>% 
    pivot_longer(-algorithms)
```

```{r}
cor_r_p_KScohort <- merge(
  rbind(
  r_KS %>% mutate(p = p_KS$value,
                  p_ = case_when(p < 0.05 ~ "*" , .default = "")) %>% 
    mutate(group = "KS"),
  
  r_M %>% mutate(p = p_M$value,
                 p_ = case_when(p < 0.05 ~ "*" , .default = "")) %>% 
    mutate(group = "Male")),
  traits_KS_df,
  by.x = "name", by.y = "traits")
  
cor_r_p_KScohort %>% 
ggplot(aes(factor(name, levels = order_cor_KS),
           factor(algorithms, levels = c(vars)),
           fill = value)) + 
  geom_tile() + 
  geom_text(aes(label = ifelse(p < 0.001, "<0.001",round(p, digits = 3)), color = p_), size = 1.5) + 
  scale_color_manual(values = c("grey","black")) + 
  facet_grid(rows = vars(factor(group, levels = c("KS","Male"))),
             cols = vars(trait_groups),
             scales = "free",
             space = "free") +
  theme_bw() +
  scale_fill_gradient2(mid = "white", high = col[4], low = col[3]) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5),
        legend.position = "right") +
  labs(x = "", y = "", title = "") -> cor_p_KS
cor_p_KS

```

```{r}
ggarrange(cor_p_TS,
          cor_p_KS,
          ncol = 2, nrow = 1,
          labels = c("A","B"),
          widths = c(1.75,1.25)) -> S1
S1

ggsave(S1, file = paste0(path_plots, "SupF1.pdf"),
                       width=12,
                       height = 12)
```

## Table 1

```{r}

# Is there a difference in cAge? 
emmeans(lm(cAge ~ group, data=df %>% filter(cohort == "TS cohort")), 
  pairwise ~ group)

emmeans(lm(cAge ~ group, data=df %>% filter(cohort == "KS cohort")), 
  pairwise ~ group)

t.test(cAge ~ group, data=df %>% filter(cohort == "TS cohort"))

sd((df %>% filter(group == "Male"))$cAge)
mean((df %>% filter(group == "TS"))$cAge)
```

```{r}
# In this section, replace the first term with whatever you are investigating 

# ageAcc
emmeans(lm(ageAcc ~ group, data = acc_df_mean %>% filter(cohort == "TS cohort")), pairwise ~ group)

((emmeans(lm(ageAcc.BLUP ~ group, data = df %>% filter(cohort == "TS cohort")), pairwise ~ group))$contrast %>% as.data.frame())$p.value

# ageAcc2
emmeans(lm(ageAcc2 ~ group, data = acc_df_mean %>% filter(cohort == "TS cohort")), pairwise ~ group)

((emmeans(lm(ageAcc2.BLUP ~ group, data = df %>% filter(cohort == "TS cohort")), pairwise ~ group))$contrast %>% as.data.frame())$p.value


# ageAcc3
emmeans(lm(ageAcc3 ~ group, data = acc_df_mean %>% filter(cohort == "TS cohort")), pairwise ~ group)

((emmeans(lm(ageAcc3.BLUP ~ group, data = df %>% filter(cohort == "TS cohort")), pairwise ~ group))$contrast %>% as.data.frame())$p.value


```

```{r}
# ageAcc
emmeans(lm(ageAcc ~ group, data = acc_df_mean %>% filter(cohort == "KS cohort")), pairwise ~ group)

((emmeans(lm(ageAcc.BLUP ~ group, data = df %>% filter(cohort == "KS cohort")), pairwise ~ group))$contrast %>% as.data.frame())$p.value

# ageAcc2
emmeans(lm(ageAcc2 ~ group, data = acc_df_mean %>% filter(cohort == "KS cohort")), pairwise ~ group)

((emmeans(lm(ageAcc2.BLUP ~ group, data = df %>% filter(cohort == "KS cohort")), pairwise ~ group))$contrast %>% as.data.frame())$p.value

# ageAcc3
emmeans(lm(ageAcc3 ~ group, data = acc_df_mean %>% filter(cohort == "KS cohort")), pairwise ~ group)

((emmeans(lm(ageAcc3.BLUP ~ group, data = df %>% filter(cohort == "KS cohort")), pairwise ~ group))$contrast %>% as.data.frame())$p.value

```

```{r}

emmeans(lm(DNAmPhenoAge ~ group,  data = pheno_df %>% filter(cohort == "TS cohort") ), pairwise ~ group)
emmeans(lm(DNAmPhenoAge ~ cAge + group,  data = pheno_df %>% filter(cohort == "TS cohort") ), pairwise ~ group)

emmeans(lm(DNAmPhenoAge ~ group,  data = pheno_df %>% filter(cohort == "KS cohort") ), pairwise ~ group)
emmeans(lm(DNAmPhenoAge ~ cAge + group,  data = pheno_df %>% filter(cohort == "KS cohort") ), pairwise ~ group)

emmeans(lm(DNAmGrimAge ~ group,  data = grim_df %>% filter(cohort == "TS cohort") ), pairwise ~ group)
emmeans(lm(DNAmGrimAge ~ cAge + group,  data = grim_df %>% filter(cohort == "TS cohort") ), pairwise ~ group)

emmeans(lm(DNAmGrimAge ~ group,  data = grim_df %>% filter(cohort == "KS cohort") ), pairwise ~ group)
emmeans(lm(DNAmGrimAge ~ cAge + group,  data = grim_df %>% filter(cohort == "KS cohort") ), pairwise ~ group)

emmeans(lm(TL ~ group,  data = TL_df %>% filter(cohort == "TS cohort") ), pairwise ~ group)
emmeans(lm(TL ~ cAge + group,  data = TL_df %>% filter(cohort == "TS cohort") ), pairwise ~ group)

emmeans(lm(TL ~ group,  data = TL_df %>% filter(cohort == "KS cohort") ), pairwise ~ group)
emmeans(lm(TL ~ cAge + group,  data = TL_df %>% filter(cohort == "KS cohort") ), pairwise ~ group)

emmeans(lm(DunedinPACE ~ group, data = pace_df %>% filter(cohort == "TS cohort") ), pairwise ~ group)
emmeans(lm(DunedinPACE ~ group, data = pace_df %>% filter(cohort == "KS cohort") ), pairwise ~ group)

```

## Table 2

```{r}

summary(lm(DNAmAge ~ cAge*group, 
           data = DNAmAge_mean_TS))

summary(lm(DNAmAge ~ cAge*group, 
           data = DNAmAge_mean_KS))

summary(lm(BLUP ~ cAge*group, 
           data = pheno_df %>% 
  filter(cohort == "TS cohort") ))

summary(lm(BLUP ~ cAge*group, 
           data = pheno_df %>% 
  filter(cohort == "KS cohort") ))

summary(lm(DNAmPhenoAge ~ cAge*group, 
           data = pheno_df %>% 
  filter(cohort == "TS cohort") ))

summary(lm(DNAmPhenoAge ~ cAge*factor(group, levels = c("Male","KS")), 
           data = pheno_df %>% 
  filter(cohort == "KS cohort") ))

summary(lm(DNAmGrimAge ~ cAge*factor(group, levels = c("Female","TS")), 
           data = grim_df %>% 
  filter(cohort == "TS cohort") ))

summary(lm(DNAmGrimAge ~ cAge*factor(group, levels = c("Male","KS")), 
           data = grim_df %>% 
  filter(cohort == "KS cohort") ))

summary(lm(TL ~ cAge*group, 
           data = TL_df %>% 
  filter(cohort == "TS cohort") ))

summary(lm(TL ~ cAge*factor(group, levels = c("Male", "KS")), 
           data = TL_df %>% 
  filter(cohort == "KS cohort") ))

summary(lm(DunedinPACE ~ cAge*factor(group, levels = c("Female", "TS")), 
           data = pace_df %>% filter(cohort == "TS cohort")))

summary(lm(DunedinPACE ~ cAge*factor(group, levels = c("Male","KS")), 
           data = pace_df %>% filter(cohort == "KS cohort")))
```

## Table S1

```{r}
t.test(`Weight (kg)` ~ group, data=CLIN)
round(range(na.omit(CLIN %>% filter(group == "TS") %>% select(`HbA1c (mmol/mol)`))),2)
round(range(na.omit(CLIN %>% filter(group == "Female") %>% select(`HbA1c (mmol/mol)`))),2)


t.test(`hemoglobin` ~ Group, data=CLIN_KS)
range(na.omit(CLIN_KS %>% filter(Group == "KS") %>% select(hemoglobin)))
range(na.omit(CLIN_KS %>% filter(Group == "Male") %>% select(hemoglobin)))

```
